\section{Related Work} \label{sec:related_work}

\paragraph{Transform coding}
Transform coding is a widely used approach in lossy image compression, leveraging mathematical transforms to decorrelate pixel values and represent image data more compactly. One of the earliest and most influential methods is the discrete cosine transform (DCT) \cite{ahmed1974discrete}, used in JPEG \cite{wallace1991jpeg}, which converts image data into the frequency domain, prioritizing lower frequencies to retain perceptually significant information. The discrete wavelet transform (DWT) \cite{antonini1992image}, used in JPEG 2000 \cite{skodras2001jpeg}, offers improved performance by capturing both frequency and location information, leading to better handling of edges and textures \cite{shapiro1993embedded}. More recently, the WebP \cite{google2011webp} and HEIF \cite{lainema2016hevc, hannuksela2015high} formats combine DCT and intra-frame prediction to achieve superior compression and quality compared to JPEG.

\paragraph{Low-rank techniques}
Low-rank approximation can provide a compact representation by decomposing image data into smaller components. Notably, truncated singular value decomposition (tSVD) is a classical technique that decomposes images into singular values and vectors, retaining only the most significant components to achieve compression \cite{andrews1976singular, prasantha2007image}. \textcite{hou2015sparse} proposed sparse low-rank matrix approximation (SLRMA) for data compression, which is able to explore both the intra- and inter-coherence of data samples simultaneously from the perspective of optimization and transformation.
More recently, \textcite{yuan2020image} introduced a graph-based low-rank regularization to reduce compression artifacts near block boundaries at low bit rates.

\paragraph{Integer Matrix Factorization}
There are applications where meaningful representation of data as discrete factor matrices is crucial. While typical low-rank techniques like SVD and nonnegative matrix factorization (NMF) are inappropriate for such applications, integer matrix factorization (IMF) ensures the integrality of factors to achieve this goal. \textcite{lin2005integer} investigates IMF to effectively handle discrete data matrices for cluster analysis and pattern discovery. \textcite{dong2018integer} introduce an alternative least squares method for IMF, verifying its effectiveness with some data mining applications. A closely related topic is boolean matrix factorization (BMF), where factors are constrained to binary matrices. BMF has been applied to some data mining \cite{miettinen2008discrete} and machine learning \cite{ravanbakhsh2016boolean} problems. However, to the best of our knowledge, IMF has not yet been explored in image compression. Therefore, this work investigates IMF for image compression and argues that it can serve as a powerful tool for this purpose.
