\begin{abstract}
Image compression is a fundamental task in image processing, essential for the efficient transmission and storage. Most widely used methods, like JPEG, use transformations in the continuous domain, such as discrete cosine transform (DCT), which necessitates carefully designed quantization techniques. Another promising approach relies on low-rank approximation techniques, such as singular value decomposition (SVD), which again requires a quantization layer. However, the compression performance of this approach is more prone to quantization errors compared to JPEG. Therefore, we introduce a variant of integer matrix factorization (IMF) that forms the basis for our quantization-free lossy image compression technique. IMF provides a low-rank representation of the image data as a product of two smaller factor matrices with bounded integer elements, thereby eliminating the need for quantization. Moreover, the reshaped factor matrices can be treated as grayscale images. Hence, any standard lossless image compression method can be seamlessly integrated into the proposed IMF encoder. We propose an efficient, provably convergent iterative algorithm for IMF using a block coordinate descent (BCD) scheme, where each column of a factor matrix is updated using a closed-form solution at a time. The experiments show that our IMF method outperforms JPEG, achieving over 4 dB and 7 dB improvement of PSNR in image quality at 0.15 bits per pixel (bpp) on Kodak and CLIC datasets, respectively. We also evaluated methods in preserving the visual semantic information by assessing a ResNet50 classifier pre-trained on ImageNet. The input images were compressed by different methods before being fed into the classifier. The results suggest that IMF leads to more than a 25\% relative improvement in top-1 accuracy over JPEG at the low bitrate of 23 bpp.
\end{abstract}